Income Disparity and Its Influence on Higher Education Affordability and Access

### 2024-02-18 starter_folder-main 4/paper/paper.qmd
updated tables

>>>>

```{r}
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)

# Set the file path
file_path <- "/Users/michaelfang/Downloads/Figure_1_data.xlsx"

## Fig 10 is public college net prices
## Fig 11 is private college net prices 
public_college_netprice <- read_excel(file_path, sheet = "Fig 10", range = cell_cols("B:V"))
private_college_netprice <- read_excel(file_path, sheet = "Fig 11", range = cell_cols("B:V"))

## Fig 2013_12 public four year total grants by income and year
## Fig 2013_13 private four year total grants by income and year

public_college_grants <- read_excel(file_path, sheet = "Fig 2013_12", range = cell_cols("B:T"))
private_college_grants <- read_excel(file_path, sheet = "Fig 2013_13", range = cell_cols("B:T"))

# modifying column names for better access
colnames(public_college_grants) <- c("1999_low", "2003_low", "2007_low", "2011_low", "empty1", "1999_sec", "2003_sec", "2007_sec", "2011_sec", "empty2", "1999_third", "2003_third", "2007_third", "2011_third", "empty3", "1999_high", "2003_high", "2007_high", "2011_high", "empty4")

# modifying column names for better access
colnames(private_college_grants) <- c("1999_low", "2003_low", "2007_low", "2011_low", "empty1", "1999_sec", "2003_sec", "2007_sec", "2011_sec", "empty2", "1999_third", "2003_third", "2007_third", "2011_third", "empty3", "1999_high", "2003_high", "2007_high", "2011_high", "empty4")

# creating the column names
expenditure_type <- c("Net TFRB", "Grants by Income(1)", "Grants by Income(2)", "Grants by Income(3)", "Grants by Income(4)", "Funding for Higher Education per 1000$ of Income", "Tuition Revanue", "Subsidy")

## Fig 14B about the total funding in billions
#funding_for_higher_education <- read_excel(file_path, sheet = "Fig 15B")

## Fig 15A average state and local funding for higher education per 1000$ of personal income
state_funding <- read_excel(file_path, sheet = "Fig 15A")
colnames(state_funding) <- c('Year','Funding Per 1000 Income')

## Fig 17A public 4 year of subsidies and revenues in different years
public_rev_subs <- read_excel(file_path, sheet = "Fig 17A")
colnames(public_rev_subs) <- c("Institution Type","Year", "Net Tuition Revenue", "Subsidy", "Education Related Expenditures")

## Fig 17B private 4 year of subsidies and revenues in different years
private_rev_subs <- read_excel(file_path, sheet = "Fig 17B")
colnames(private_rev_subs) <- c("Institution Type","Year", "Net Tuition Revenue", "Subsidy", "Education Related Expenditures")

```
\newpage

## Public Four Year College Data

```{r}
#| message: false
#| echo: false
#| label: tbl-public_data
#| tbl-cap: Public Four Year College Data

# this is to shorten the name allowing cleaner code and set all items to numeric
pcn <- public_college_netprice
pcg <- public_college_grants

# rounding numbers to get rid of long decimals
a <- round(as.numeric(pcg$"1999_low"[13]), 0)
a1 <- round(as.numeric(pcg$"1999_sec"[13]), 0)
a2 <-round(as.numeric(pcg$"1999_third"[13]), 0)
a3 <-round(as.numeric(pcg$"1999_high"[13]), 0)
a4 <-round(as.numeric(state_funding$`Funding Per 1000 Income`[1]), 3)

# creating column of data1
pub_earliest_data <-c(pcn$"96-97"[4], a, a1, a2, a3, a4, public_rev_subs$"Net Tuition Revenue"[2], public_rev_subs$"Subsidy"[2])

b <- round(as.numeric(pcg$"2011_low"[13]), 0)
b1 <- round(as.numeric(pcg$"2011_sec"[13]), 0)
b2 <- round(as.numeric(pcg$"2011_third"[13]), 0)
b3 <- round(as.numeric(pcg$"2011_high"[13]), 0)
b4 <- round(as.numeric(state_funding$`Funding Per 1000 Income`[31]), 3)

pub_latest_data <-c(public_college_netprice$"16-17"[4],  b, b1, b2, b3, b4, public_rev_subs$"Net Tuition Revenue"[4], public_rev_subs$"Subsidy"[4])

pub_percent_change <-c("62.77%", "68.21%", "80.53%", "69.74%", "78.52%", "-28.31%", "47.35%", "-8.5%")

pub_year_frame <-c("1996 to 2016", "1999 to 2011","1999 to 2011", "1999 to 2011","1999 to 2011","1984 to 2014", "2003 to 2013", "2003 to 2013")

# create frame for public data
public_data <- tibble("Type" = expenditure_type, "Data_1" = pub_earliest_data, "Data_2" = pub_latest_data, "Percent_Change" = pub_percent_change, "Time_Frame" = pub_year_frame)

# create the table
public_data_table <-
head (public_data, 8) |>
  kable(
  booktabs = TRUE
  )

public_data_table
```
## Private Four Year College Data

```{r}
#| message: false
#| echo: false
#| label: tbl-private_data
#| tbl-cap: Private Four Year College Data

prcn <- private_college_netprice
prcg <- private_college_grants

# rounding data and making sure that number is not too long
c <- round(as.numeric(prcg$"1999_low"[4]), 0)
c1 <- round(as.numeric(prcg$"1999_sec"[4]), 0)
c2 <-round(as.numeric(prcg$"1999_third"[4]), 0)
c3 <-round(as.numeric(prcg$"1999_high"[4]), 0)
c4 <-round(as.numeric(state_funding$`Funding Per 1000 Income`[1]), 3)

# creating data 1 column
pri_earliest_data <-c(prcn$"96-97"[4],c, c1, c2, c3, c4, private_rev_subs$"Net Tuition Revenue"[2], private_rev_subs$"Subsidy"[2])

# rounding data and making sure that number is not too long
d <- round(as.numeric(prcg$"2011_low"[4]), 0)
d1 <- round(as.numeric(prcg$"2011_sec"[4]), 0)
d2 <-round(as.numeric(prcg$"2011_third"[4]), 0)
d3 <-round(as.numeric(prcg$"2011_high"[4]), 0)
d4 <-round(as.numeric(state_funding$`Funding Per 1000 Income`[31]), 3)

pri_latest_data <-c(prcn$"16-17"[4], d, d1,d2, d3, d4, private_rev_subs$"Net Tuition Revenue"[4], private_rev_subs$"Subsidy"[4])

pri_percent_change <-c("30.27%", "120%", "80.84%", "28.30%", "97.36%", "-28.31%", "17.62%", "28.86%")

pri_year_frame <-c("1996 to 2016", "1999 to 2011","1999 to 2011", "1999 to 2011","1999 to 2011","1984 to 2014", "2003 to 2013", "2003 to 2013")

# create frame for private data
private_data <- data.frame("Type" = expenditure_type, "Data_1" = pri_earliest_data, "Data_2" = pri_latest_data, "Percent_Change" = pri_percent_change, "Year_Frame" = pri_year_frame)

# create the table
private_data_table <-
head (private_data, 8) |>
  kable(
  booktabs = TRUE
  )
private_data_table

```
